---
title: "label transfer"
output: html_notebook
---

# Set some user parameters

```{r}
sampleID <- "assayID"
outdir <- "/path/to/outdir/"
df <- readRDS("/path/to/your.RDS") #the RDS output of SeuratSpatialAnalysisPipeline.Rmd
refID <- "refID" #used to save your files - e.g. Lake or Liao
ref <- readRDS("/path/to/reference.RDS") #pre-processed Seurat object of reference data, e.g. Lake or Liao

#Note: this code assumes the metadata of interest you are "transferring" is in a column called "annotation_simple". If this is not the case, cntl+f for "annotation_simple" and change to the column name you wish to use for annotation
```

# load the data

```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
```

# Run label transfer. 

We're going to match up our query data (Visium) with the reference data and make connections/"anchors" between cells that can be considered to be the same type. Then we can use the annotation of each reference cell to annotate the corresponding query cell/s. 

The query dataset can ONLY be given annotations that exist in the reference set, and will be given a label even if there are no "true" hits. For example, if you annotate a kidney dataset with a neuronal reference, every cell will still get a label. However, we expect that the probability scores associated with each annotation will be low.

Therefore, rather than accept all the labels we are given, we also add an additional filtering step where any annotation with a prediction score less than 0.6 (60%) is relabeled as "unknown".

```{r}
anchors <- FindTransferAnchors(reference = ref, query = df, project.query = T)
predictions <- TransferData(anchorset = anchors, refdata = ref$annotation_simple)

predictions_sml <- predictions[,c("predicted.id", "prediction.score.max")]
colnames(predictions_sml)[colnames(predictions_sml)=="predicted.id"] <- paste0(refID, "_abbrev_predicted.id")
colnames(predictions_sml)[colnames(predictions_sml)=="prediction.score.max"] <- paste0(refID, "_abbrev_prediction.score.max")
df <- AddMetaData(df, metadata = predictions_sml)
# and now rename the annotations with scores that are too low
predictions_sml <- predictions[,c("predicted.id", "prediction.score.max")]
predictions_sml$predicted.id <- ifelse(predictions_sml$prediction.score.max < 0.6, NA, predictions_sml$predicted.id)
predictions_sml <- predictions_sml[,1, drop = FALSE]
colnames(predictions_sml)[colnames(predictions_sml)=="predicted.id"] <- paste0(refID, "_abbrev_predicted.id_hiscore")
# apply metadata
df <- AddMetaData(df, metadata = predictions_sml)
write.table(predictions_sml, file = paste0(outdir, sampleID, ".vs.", refID, "_labelTransferMetadata.txt"))
saveRDS(df, file = paste0(outdir, sampleID, ".vs.", refID, "_labelTransferResults.RDS"))
```

```{r}
# plot showing distribution of prediction scores
SpatialFeaturePlot(df, features = paste0(refID, "_abbrev_prediction.score.max"), alpha = c(0.5, 1))
ggsave(file = paste0(outdir, sampleID, ".vs.", refID, "_labelTransfer_SpatialPredictionScores.pdf"))
```


```{r fig.height=12, fig.width=12}
# run for all labels, regardless of score
DimPlot(df, group.by = paste0(refID, "_abbrev_predicted.id"), pt.size = 1)
ggsave(file = paste0(outdir, sampleID, ".vs.", refID, "_labelTransfer_AllAnnotations_DimPlot.pdf"))


SpatialDimPlot(df, group.by = paste0(refID, "_abbrev_predicted.id"))
ggsave(file = paste0(outdir, sampleID, ".vs.", refID, "_labelTransfer_AllAnnotations_SpatialPlot.pdf"))

# run for all labels with >60% confidence
DimPlot(df, group.by = paste0(refID, "_abbrev_predicted.id_hiscore"), pt.size = 1)
ggsave(file = paste0(outdir, sampleID, ".vs.", refID, "_labelTransfer_HiScoreAnnotations_DimPlot.pdf"))
SpatialDimPlot(df, group.by = paste0(refID, "_abbrev_predicted.id_hiscore"))
ggsave(file = paste0(outdir, sampleID, ".vs.", refID,  "_labelTransfer_HiScoreAnnotations_SpatialPlot.pdf"))
```

# annotation counts

```{r}
table_hi <- as.data.frame(table(df[[paste0(refID, "_abbrev_predicted.id_hiscore")]]))
ggplot(table_hi, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity") +
  coord_flip()
ggsave(file = paste0(outdir, sampleID, ".vs.", refID, "_abbrev_hiscore_counts.pdf"))

table_all <- as.data.frame(table(df[[paste0(refID, "_abbrev_predicted.id")]]))
ggplot(table_all, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity") +
  coord_flip()
ggsave(file = paste0(outdir, sampleID, ".vs.", refID, "_abbrev_allscores_counts.pdf"))
```

# Marker genes for transferred labels (high score)

```{r}
Idents(df) <- paste0(refID, "_abbrev_predicted.id_hiscore")
hiscoremarkers <- FindAllMarkers(df, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
hitop10 <- hiscoremarkers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
write.table(hiscoremarkers, file = paste0(outdir, sampleID, ".vs.", refID, "_allmarkers_highscore.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(hitop10, file = paste0(outdir, sampleID, ".vs.", refID, "_top10markers_highscore.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

# Marker genes for clusters (all hits)

```{r}
Idents(df) <- paste0(refID, "_abbrev_predicted.id")
allmarkers <- FindAllMarkers(df, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
alltop10 <- allmarkers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
write.table(allmarkers, file = paste0(outdir, sampleID, ".vs.", refID, "_allmarkers_anyscore.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(alltop10, file = paste0(outdir, sampleID, ".vs.", refID, "_top10markers_anyscore.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

